Stochastic Variance Reduced Policy Gradient

### 2018-01-15 svrpg.tex
written main theorem

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%%%%%%

@@

\section{Algorithm}
In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment. We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{itemize}
	\item The objective $J(\vtheta)$ is typically non-convex;
	\item The dataset $\mathcal{T}$ can be infinite;
	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
\end{itemize}
To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?]. Because of the infinite dataset, we can only rely on an estimate of the full gradient. Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories. We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward. The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
\begin{align*}
	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
\end{align*}
where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from the snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. We immediately have the following (proof in Appendix \ref{app:proofs}):

\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
The SVRPG gradient estimator is unbiased:
\[
\mathop{E}_{\tau\sim\pi(\cdot\mid\tilde{\vtheta})}\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
\]
\end{restatable}

@@

@@

\end{document}

>>>>

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla_{#1}\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla_{#1}J(\vtheta)}
\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%%%%%%

@@

\section{Algorithm}
In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment. We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{itemize}
	\item The objective $J(\vtheta)$ is typically non-convex;
	\item The dataset $\mathcal{T}$ can be infinite;
	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
\end{itemize}
To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?]. Because of the infinite dataset, we can only rely on an estimate of the full gradient. Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories. We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward. The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
\begin{align*}
	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
\end{align*}
where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. We immediately have the following (proof in Appendix \ref{app:proofs}):

\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
The SVRPG gradient estimator is unbiased:
\[
\mathop{E}_{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}),\tau_i\sim\pi(\cdot\mid\vtheta)}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{}.
\]
\end{restatable}

@@

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
The parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, b>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
\[
	\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} \leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{Z}{N},
\]
where
\begin{align*}
	Z &= b\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
	&+ \max_{s,t}\left|\Tr\Cov\left[
		\gradApp{\tilde{\vtheta^s}}{1},
		c_{t+1}\vtheta_t^{s+1} + \alpha_t\nabla J(\vtheta_t^{s+1})\right]\right|
\end{align*}
\end{restatable}

@@

\convergence*
\begin{proof}
	...
\end{proof}
\end{document}
